---
title: "Image analysis for Hannah Bailey - preliminary analysis"
subtitle: ""
author: "Dan Kelpsch"
format:
  html:
    self-contained: true
editor: visual
---

# Background

Hannah has been working on various drug treatments and found that they alter lysosome and pRab10+ vesicles morphology and localization. She has been interested in trying to quantify this phenotype, but it has proven difficult. We have generated segmentation algorithms to generate masks of the cells in her images, lysosomes, and pRab10+ vesicles and extracted associated features and colocalization metrics. Here is the quantification of some of those features.

# Load packages

```{r}
#| include: false
#| label: Load necessary packages
list.of.packages <- c("tidyverse")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, library, character.only = TRUE)

##round_any() is a function in the package plyr, but plyr has some 
##compatibility issues with the tidyverse, thus just writing in the 
##function manually
round_any = function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
```

# Load and process data

```{r}
#| label: Load data
lysosomes <- read_csv("Hannah_RepresentativeImages/pipeline_outputs/lys_objects_all.csv")
prab <- read_csv("Hannah_RepresentativeImages/pipeline_outputs/prab_objects_all.csv")
coloc <- read_csv("Hannah_RepresentativeImages/pipeline_outputs/cells_coloc_all.csv")
samplesheet <- read_csv("Hannah_RepresentativeImages/hb_samplesheet.csv")

dir.create("./Output_graphs", showWarnings = FALSE)
```

## Align data with sample sheet

```{r}
#| label: Data processing
lysosomes <- lysosomes %>% 
  left_join(samplesheet)

average_lysosomes <- lysosomes %>% 
  group_by(cell_id, filename, Treatment) %>% 
  summarize(mean_area = mean(area),
            mean_fint = mean(mean_intensity))

prab <- prab %>% 
  left_join(samplesheet)

average_prab <- prab %>% 
  group_by(cell_id, filename, Treatment) %>% 
  summarize(mean_area = mean(area),
            mean_fint = mean(mean_intensity))

coloc <- coloc %>%
  left_join(samplesheet)
```

# Lysosome analysis

## Count

```{r}
#| label: Lysosome count
counts_per_image <- lysosomes %>%
  group_by(Treatment, filename, cell_id) %>%
  summarise(count = n(), .groups = "drop")

y_lim <- counts_per_image %>% 
  #filter(Outlier == "Normal") %>%
  pull(count) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 100, ceiling)

ggplot(data = counts_per_image,
       aes(x = Treatment,
           y = count)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/6)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Lysosome count (per cell)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "lysosome_count.png",  width = 6, height = 4, path = "./Output_graphs/")
```

## Size

```{r}
#| label: Lysosome size
y_lim <- lysosomes %>% 
  #filter(Outlier == "Normal") %>%
  pull(area) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 100, ceiling)

ggplot(data = lysosomes,
       aes(x = Treatment,
           y = area)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  #geom_jitter(width = 0.2, alpha = 0.05) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/7)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Lysosome size (px)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "lysosome_size.png",  width = 6, height = 4, path = "./Output_graphs/")

y_lim <- average_lysosomes %>% 
  #filter(Outlier == "Normal") %>%
  pull(mean_area) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 10, ceiling)

ggplot(data = average_lysosomes,
       aes(x = Treatment,
           y = mean_area)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Average lysosome size (per cell)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "lysosome_average_size.png",  width = 6, height = 4, path = "./Output_graphs/")
```

## Intensity

```{r}
#| label: Lysosome intensity
y_lim <- lysosomes %>% 
  #filter(Outlier == "Normal") %>%
  pull(mean_intensity) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 100, ceiling)

ggplot(data = lysosomes,
       aes(x = Treatment,
           y = mean_intensity)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  #geom_jitter(width = 0.2, alpha = 0.05) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/7)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Mean fluorescent intensity (per lysosome mask)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "lysosome_fintensity.png",  width = 6, height = 4, path = "./Output_graphs/")

y_lim <- average_lysosomes %>% 
  #filter(Outlier == "Normal") %>%
  pull(mean_fint) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 10, ceiling)

ggplot(data = average_lysosomes,
       aes(x = Treatment,
           y = mean_fint)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Average lysosome size (per cell)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 
```

# pRab10+ vesicle analysis

## Count

```{r}
#| label: pRab10+ vesicle count
counts_per_image <- prab %>%
  group_by(Treatment, filename, cell_id) %>%
  summarise(count = n(), .groups = "drop")

y_lim <- counts_per_image %>% 
  #filter(Outlier == "Normal") %>%
  pull(count) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 100, ceiling)

ggplot(data = counts_per_image,
       aes(x = Treatment,
           y = count)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "pRab+ object count (per cell)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "prab_count.png",  width = 6, height = 4, path = "./Output_graphs/")
```

## Size

```{r}
#| label: pRab10+ vesicle size
y_lim <- prab %>% 
  #filter(Outlier == "Normal") %>%
  pull(area) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 100, ceiling)

ggplot(data = prab,
       aes(x = Treatment,
           y = area)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  #geom_jitter(width = 0.2, alpha = 0.05) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "pRab+ object size (px)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "prab_size.png",  width = 6, height = 4, path = "./Output_graphs/")

y_lim <- average_prab %>% 
  #filter(Outlier == "Normal") %>%
  pull(mean_area) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 10, ceiling)

ggplot(data = average_prab,
       aes(x = Treatment,
           y = mean_area)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Average pRab+ object size (per cell)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "prab_average_size.png",  width = 6, height = 4, path = "./Output_graphs/")
```

## Intensity

```{r}
#| label: pRab10+ vesicle intensity
y_lim <- prab %>% 
  #filter(Outlier == "Normal") %>%
  pull(mean_intensity) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 500, ceiling)

ggplot(data = prab,
       aes(x = Treatment,
           y = mean_intensity)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  #geom_jitter(width = 0.2, alpha = 0.05) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Mean fluorescent intensity (per pRab+ object mask)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "prab_fintensity.png",  width = 6, height = 4, path = "./Output_graphs/")

y_lim <- average_prab %>% 
  #filter(Outlier == "Normal") %>%
  pull(mean_fint) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 10, ceiling)

ggplot(data = average_prab,
       aes(x = Treatment,
           y = mean_fint)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Average pRab+ object fluorescent intensity (per cell)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "prab_average_fintensity.png",  width = 6, height = 4, path = "./Output_graphs/")
```

# Colocalization analysis

## Pearson's

```{r}
#| label: Colocalization Pearson's R
y_lim <- coloc %>% 
  #filter(Outlier == "Normal") %>%
  pull(pearson_r) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 1, ceiling)

ggplot(data = coloc,
       aes(x = Treatment,
           y = pearson_r)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(-0.1,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Colocalization (Pearson's R)") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "coloc_pearsons.png",  width = 6, height = 4, path = "./Output_graphs/")
```

## Mander's

```{r}
#| label: Colocalization Mander's M1
y_lim <- coloc %>% 
  #filter(Outlier == "Normal") %>%
  pull(manders_M1_lys_in_prab) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 1, ceiling)

ggplot(data = coloc,
       aes(x = Treatment,
           y = manders_M1_lys_in_prab)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(-0.1,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Colocalization (Manders (lys in prab))") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "coloc_M1_lys_in_prab.png",  width = 6, height = 4, path = "./Output_graphs/")
```

```{r}
#| label: Colocalization Mander's M2
y_lim <- coloc %>% 
  #filter(Outlier == "Normal") %>%
  pull(manders_M2_prab_in_lys) %>% 
  max(., na.rm = TRUE) %>% 
  round_any(., 1, ceiling)

ggplot(data = coloc,
       aes(x = Treatment,
           y = manders_M2_prab_in_lys)) +
  geom_violin(fill = "skyblue",
              alpha = 0.7,
              scale = "width") +
  geom_boxplot(width = 0.5,
               outlier.shape = NA,
               alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::comma, limits = c(0,y_lim), #c(lower limit, upper limit)
    breaks = seq(0,y_lim,y_lim/10)) +  #seq(lower limit, upper limit, tick frequency)
  labs(x = "Drug treatment",
       y = "Colocalization (Manders (prab in lys))") +
  theme(
    axis.text.x = element_text(face = 'bold', angle = 45, hjust = 1), 
    axis.title.x = element_text(face = 'bold'),
    axis.text.y = element_text(face = 'bold'), 
    axis.title.y = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.text = element_text(face = 'bold'),
    strip.text.x = element_text(face = 'bold')) 

ggsave(filename = "coloc_M2_prab_in_lys.png",  width = 6, height = 4, path = "./Output_graphs/")
```
